### 2022-07-14 thesis/staticalg.tex
Adding evaluation tables and plots
First draft of thesis more or less completed
slight modifications and corrections in prelims

We will end with an outline of the static algorithm first investigated in works \cite{dang2012reachability, dreossi2016parallelotope}. As mentioned in the previous section, the building block of the reachability algorithm relies on approximate solutions to a non-linear optimization problem over the unitbox domain. Consider a nonlinear function $h : \reals^n \rightarrow \reals$. The most general form of this optimization problem can be expressed as:
%
\begin{eqnarray}
  \max ~ h(x) \label{eq:maxsup}\\
  s.t. ~~ x \in [0,1]^{n}.\nonumber
\end{eqnarray}

@@

We now proceed to formally describe the static algorithm.
%
First, a small remark on the template matrix of the parallelotopes $P_i$ contained in some bundle $Q$. It is possible that some of the parallelotopes share the same template matrix directions.
%
 In other words, for $P_i = \langle \Lambda^{P_i}, c^{P_i} \rangle, P_j = \langle \Lambda^{P_j}, c^{P_j} \rangle$ such that $P_i, P_j \in \ptopeset{Q}$, there could exist some $k$ such that $\Lambda^{P_i}_k = \Lambda^{P_j}_k$ as row vectors.
%
Thus, a more compact method of encoding the bundle is by taking the \emph{distinct} template directions as rows of a new template matrix $\Lambda^Q$ along with its corresponding offset vector $c^Q$.
%
To distinguish between the distinct parallelotopes contained in the bundle, we add a new matrix called $\T^Q \in \mathbb{N}^{p \times n}$ such that $\T^Q_i$ is a vector of row indices of $\Lambda^Q$ which specify the template directions defining parallelotope $P_i \in \ptopeset{Q}$.
%
\begin{remark}
If none of the paralleotopes $P_i \in \ptopeset{Q}$ share common template directions, then $\Lambda^Q$ will simply be the template direction matrices $\{\Lambda^P\}_{P \in \ptopeset{Q}}$ concatented along their rows. This will generally be the matrix generated by the dynamic algorithm we will outline in a future section.
\end{remark}
%
\begin{example}

@@

%
% We would like to compute the image of the parallelotope $P$ with respect to the nonlinear transformation $f(x)$.
%
% That is, we would like to compute $P'$ such that $\forall x \in P, f(x) \in P'$.
%
% Given that current techniques use static template directions, the parallelotope $P'$ also uses the template directions $\T^{P}$.
%
To compute the upper and lower bounds of each template direction $\Lambda_{i} f(x)$, for all $x \in P$, we perform the following optimization.
%
\begin{eqnarray}
  \max ~ \Lambda_i^{P} \cdot f(x) \label{eq:maxf}\\
  s.t. ~~ x \in P.\nonumber
\end{eqnarray}
%
\noindent Note that $\Lambda_i^{P} \cdot f(x)$ is a dot product between the row vector $\Lambda_i^{P}$ and the component-wise dynamics of $f(x)$. This is similar to the method of computing support functions over convex sets \cite{boyd2004convex}.

@@

Equation~\ref{eq:maxalpha} is a form of $\mathsf{optBox(\Lambda_{i} \cdot f)}$ over $[0,1]^n$.
%
One can compute an upper-bound to this nonlinear optimization by computing the Bernstein coefficients of $\Lambda_i \cdot f(a + \Sigma_{i=1}^{n} \alpha_i g_i)$ and taking the maximum and minimum coefficients as shown in Property~\ref{prop:bern_enclosure}.
%
Similarly, we compute the lower-bound of $\Lambda_{i}\cdot f(x)$ for all $x \in P$ by computing the upperbound of $-1 \times \Lambda_{i}\cdot f(x)$.
%

We iterate this process (i.e., computing the upper and lower bound of $\Lambda^Q_{i}\cdot f(x)$) for each parallelotope in the bundle $Q$ according to Equation \ref{eq:max_over_ptopes} and Equation \ref{eq:min_over_ptopes}).
%
% Observe that the role of the parallelotopes in the bundle is to compute an upper bound on the maximum value of $T_{i}f(x)$ by expanding the domain from $Q$ to the corresponding parallelotope $P$.
%
Therefore, the tightest upper bound on $\Lambda^Q_{i}\cdot f(x)$ over $Q$ is the least of the upper bounds computed from each of the parallelotopes.
%
A similar argument holds for lower bounds of $\Lambda^Q_{i}\cdot f(x)$ over $Q$.
%
Therefore, the image of the bundle $Q$ will be the bundle $Q'$ where the upper and lower bounds for templates directions are obtained by solving a series of nonlinear optimization problems of the form presented in Equation~\ref{eq:maxf}.

@@

%\input{static_alg_fig}

@@

\begin{example}
To include an example of a higher-dimensional non-linear system, we introduce the Phosporaley model. The Phosphoraley model describes a certain cellular regulatory system. It is captured by seven variables governed by the following discretized dynamics:
\begin{align}
  \begin{split}
      x^1_{k+1} &= x^1_k + ( -\alpha \cdot x^1_k + \beta \cdot x^3_k x^4_k)\cdot \Delta \\
      x^2_{k+1} &= x^2_k + (  \alpha\cdot  x^1_k - x^2_k)\cdot \Delta \\
      x^3_{k+1} &= x^3_k + ( x^2_k - \beta \cdot x^3_k x^4_k)\cdot \Delta \\
      x^4_{k+1} &= x^4_k + ( \beta \cdot x^5_k x^6_k - \beta \cdot x^3_k x^4_k)\cdot \Delta \\
      x^5_{k+1} &= x^5_k + ( -\beta \cdot x^5_k x^6_k + \beta \cdot x^3_k x^4_k)\cdot \Delta \\
      x^6_{k+1} &= x^6_k + ( \alpha\cdot  x^7_k - \beta \cdot x^5_k x^6_k)\cdot \Delta \\
      x^7_{k+1} &= x^7_k + ( -\alpha\cdot  x^7_k + \beta \cdot  x^5_k x^6_k)\cdot \Delta
  \end{split}
\end{align}
Here, we set the two parameters $\alpha,\beta$ as $\alpha=0.5$ and $\beta=5$. The discretization step is set to $\Delta =0.01$ and we propagate the reachble set for $S=300$ time steps. Additionally, the initial box is set to be $[1.00,1.01]^7$.
%
Figure \ref{fig:kaa_phos} depicts the projection of the reachable set on the first three variables $x_1, x_2, x_3$. The paralleotopes are the axis-aligned parallotope and another with the following template direction matrix:

\begin{equation}
  \Lambda = \begin{bmatrix}
            1 & 0 & 0 & 0 & 0 & 0 & 0 \\
            0 & 1 & 0 & 0 & 0 & 0 & 0 \\
            0 & 0 & 1 & 0 & 0 & 0 & 0 \\
            0 & 0 & 0 & 0 & 1 & 0 & 0 \\
            0 & 0 & 0 & 0 & 0 & 1 & 0 \\
            0 & 0 & 0 & 0 & 0 & 0 & 1 \\
            0 & 0 & 1 & 1 & 0 & 0 & 0 \\
            \end{bmatrix}
\end{equation}

\begin{figure}
\label{fig:kaa_phos}
  \hspace*{-2.3cm}
  \includegraphics[width=1.3\textwidth]{figures/PhosporlayProjOn X_1, X_2, X_3.png}
  \caption{Projection of Reachable Set of the Phosporaley model propagated 300 steps in time.}
\end{figure}
\end{example}

>>>>

We will end with an outline of the static algorithm first investigated in works \cite{dang2012reachability, dreossi2016parallelotope}. As mentioned in the previous section, the building block of the reachability algorithm relies on approximate solutions to a non-linear optimization problem over the unitbox domain. Consider a non-linear function $h : \reals^n \rightarrow \reals$. The most general form of this optimization problem can be expressed as:
%
\begin{eqnarray}
  \max ~ h(x) \label{eq:maxsup}\\
  s.t. ~~ x \in [0,1]^{n}.\nonumber
\end{eqnarray}

@@

We now proceed to formally describe the static algorithm.
%
First, a small remark on the template matrix of the parallelotopes $P_i$ contained in some bundle $Q$. It is possible that some of the parallelotopes share the same template matrix directions.
%
 In other words, for $P_i = \langle \Lambda^{P_i}, c^{P_i} \rangle, P_j = \langle \Lambda^{P_j}, c^{P_j} \rangle$ such that $P_i, P_j \in \ptopeset{Q}$, there could exist some $k$ such that $\Lambda^{P_i}_k = \Lambda^{P_j}_k$ as row vectors.
%
Thus, a more compact method of encoding the bundle is by taking the \emph{distinct} template directions as rows of a new template matrix $\Lambda^Q$ along with its corresponding offset vector $c^Q$.
%
To distinguish between the distinct parallelotopes contained in the bundle, we add a new matrix called $\T^Q \in \mathbb{N}^{p \times n}$ such that $\T^Q_i$ is a vector of row indices of $\Lambda^Q$ which specify the template directions defining parallelotope $P_i \in \ptopeset{Q}$.
%
The number of rows will thus be the number of distinct parallelotopes contained in the bundle $p = |\ptopeset{Q}|$.
%
\begin{remark}
If none of the paralleotopes $P_i \in \ptopeset{Q}$ share common template directions, then $\Lambda^Q$ will simply be the template direction matrices $\{\Lambda^P\}_{P \in \ptopeset{Q}}$ concatented along their rows. This will generally be the matrix generated by the dynamic algorithm we will outline in a future section.
\end{remark}
%
\begin{example}

@@

%
% We would like to compute the image of the parallelotope $P$ with respect to the non-linear transformation $f(x)$.
%
% That is, we would like to compute $P'$ such that $\forall x \in P, f(x) \in P'$.
%
% Given that current techniques use static template directions, the parallelotope $P'$ also uses the template directions $\T^{P}$.
%
To compute the upper and lower bounds of each template direction $\Lambda_{i} f(x)$, for all $x \in P$, we perform the following optimization.
%
\begin{eqnarray}
  \max ~ \Lambda_i^{P} \cdot f(x) \label{eq:maxf}\\
  s.t. ~~ x \in P.\nonumber
\end{eqnarray}
%
\noindent Note that $\Lambda_i^{P} \cdot f(x)$ is a dot product between the row vector $\Lambda_i^{P}$ and the component-wise dynamics of $f(x)$. This is similar to the method of computing support functions over convex sets \cite{boyd2004convex}.

@@

Equation~\ref{eq:maxalpha} is a form of $\mathsf{optBox(\Lambda_{i} \cdot f)}$ over $[0,1]^n$.
%
One can compute an upper-bound to this non-linear optimization by computing the Bernstein coefficients of $\Lambda_i \cdot f(a + \Sigma_{i=1}^{n} \alpha_i g_i)$ and taking the maximum and minimum coefficients as shown in Property~\ref{prop:bern_enclosure}.
%
Similarly, we compute the lower-bound of $\Lambda_{i}\cdot f(x)$ for all $x \in P$ by computing the upperbound of $-1 \times \Lambda_{i}\cdot f(x)$.
%

We iterate this process (i.e., computing the upper and lower bound of $\Lambda^Q_{i}\cdot f(x)$) for each parallelotope in the bundle $Q$ according to Equation \ref{eq:max_over_ptopes} and Equation \ref{eq:min_over_ptopes}).
%
% Observe that the role of the parallelotopes in the bundle is to compute an upper bound on the maximum value of $T_{i}f(x)$ by expanding the domain from $Q$ to the corresponding parallelotope $P$.
%
Therefore, the tightest upper bound on $\Lambda^Q_{i}\cdot f(x)$ over $Q$ is the least of the upper bounds computed from each of the parallelotopes.
%
A similar argument holds for lower bounds of $\Lambda^Q_{i}\cdot f(x)$ over $Q$.
%
Therefore, the image of the bundle $Q$ will be the bundle $Q'$ where the upper and lower bounds for templates directions are obtained by solving a series of non-linear optimization problems of the form presented in Equation~\ref{eq:maxf}.

@@

\input{static_alg_fig}

@@

\begin{example}

To include an example of a higher-dimensional non-linear system, we introduce the Phosporaley model. The Phosphoraley model describes a certain cellular regulatory system. It is captured by seven variables governed by the following discretized dynamics:
\begin{align}
  \begin{split}
      x^1_{k+1} &= x^1_k + ( -\alpha \cdot x^1_k + \beta \cdot x^3_k x^4_k)\cdot \Delta \\
      x^2_{k+1} &= x^2_k + (  \alpha\cdot  x^1_k - x^2_k)\cdot \Delta \\
      x^3_{k+1} &= x^3_k + ( x^2_k - \beta \cdot x^3_k x^4_k)\cdot \Delta \\
      x^4_{k+1} &= x^4_k + ( \beta \cdot x^5_k x^6_k - \beta \cdot x^3_k x^4_k)\cdot \Delta \\
      x^5_{k+1} &= x^5_k + ( -\beta \cdot x^5_k x^6_k + \beta \cdot x^3_k x^4_k)\cdot \Delta \\
      x^6_{k+1} &= x^6_k + ( \alpha\cdot  x^7_k - \beta \cdot x^5_k x^6_k)\cdot \Delta \\
      x^7_{k+1} &= x^7_k + ( -\alpha\cdot  x^7_k + \beta \cdot  x^5_k x^6_k)\cdot \Delta
  \end{split}
\end{align}
Here, we set the two parameters $\alpha,\beta$ as $\alpha=0.5$ and $\beta=5$. The discretization step is set to $\Delta =0.01$ and we propagate the reachble set for $S=300$ time steps. Additionally, the initial box is set to be $[1.00,1.01]^7 \times [-100, 100]$.
%
\begin{figure}[t!]
\label{fig:kaa_phos}
  \hspace*{-2.3cm}
  \includegraphics[width=1.3\textwidth]{figures/PhosporlayProjOn X_1, X_2, X_3.png}
  \caption{Projection of Reachable Set of the Phosporaley model propagated 300 steps in time.}
\end{figure}
%
Figure \ref{fig:kaa_phos} depicts the projection of the reachable set on the first three variables $x_1, x_2, x_3$. The relevant matrices are defined as:

\begin{equation}
  \Lambda^Q = \begin{bmatrix}
            1 & 0 & 0 & 0 & 0 & 0 & 0 \\
            0 & 1 & 0 & 0 & 0 & 0 & 0 \\
            0 & 0 & 1 & 0 & 0 & 0 & 0 \\
            0 & 0 & 0 & 1 & 0 & 0 & 0 \\
            0 & 0 & 0 & 0 & 1 & 0 & 0 \\
            0 & 0 & 0 & 0 & 0 & 1 & 0 \\
            0 & 0 & 0 & 0 & 0 & 0 & 1 \\
            0 & 0 & 1 & 1 & 0 & 0 & 0 \\
            \end{bmatrix}, \quad
  \T^Q = \begin{bmatrix}
        0 & 1 & 2 & 3 & 4 & 5 & 6 \\
        0 & 1 & 2 & 4 & 5 & 6 & 7
        \end{bmatrix}
\end{equation}
$\T^Q$ tells us that the bundle consists of the axis-aligned parallelotope (the first row $\T^Q_1$) and another diagonal parallelotope (the second row $\T^Q_2$). 
\end{example}
